\documentclass[11pt]{article}
\usepackage[a4paper,margin=25mm]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{tikz-cd}
\usepackage{hyperref}

\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}

\title{Hiroshima Responsibility Functor}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Claim}

Hiroshima placement can be modeled as a functor
\[
  F_H:\mathcal{C}\to\mathcal{R}
\]
from the ordinary corporate decision category to the Hiroshima responsibility
category.  The functor maps the ordinary composite
\[
  D\to P\to O
\]
to the responsibility composite
\[
  D_H\to P_H\to B\to E\to V\to A\to O',
\]
which factors through both Beacon $B$ and Verification $V$.

The theorem-level claim is not that psychological pressure is directly proved.
The theorem-level claim is that paths bypassing $B$ or $V$ are not admissible
under the predicate $\mathrm{Adm}_{B,V}$.

\section*{Why the Subscript $H$ Is Hiroshima}

The subscript $H$ is not a decorative symbol.  It is justified by an external
historical premise: Hiroshima is the place where a nuclear-layer failure of
admissibility is preserved as institutional world memory and is now connected
to international AI governance.

In this formalization, ``nuclear-layer'' means the technological layer at which
humanity can damage the conditions of its own continued existence.  AI is
placed in the same layer not because it is identical to nuclear technology, but
because advanced AI governance also concerns decisions whose failure modes can
scale beyond ordinary local harm.

The Hiroshima-specific premise is the conjunction of three conditions:
\[
\begin{array}{ll}
\text{(i)} & \text{a historical witness of } \neg\mathrm{Adm}_{B,V}
             \text{ at civilizational scale},\\
\text{(ii)}& \text{institutional preservation as world memory},\\
\text{(iii)}& \text{an existing AI governance anchor.}
\end{array}
\]
Condition (i) refers to the atomic bombing of Hiroshima.  Condition (ii) is
represented by institutions such as the Hiroshima Peace Memorial / Genbaku Dome
and the Hiroshima National Peace Memorial Hall for the Atomic Bomb
Victims.\footnote{See UNESCO, ``Hiroshima Peace Memorial (Genbaku Dome)'',
\url{https://whc.unesco.org/en/list/775/}; Hiroshima National Peace Memorial
Hall for the Atomic Bomb Victims,
\url{https://www.hiro-tsuitokinenkan.go.jp/en/about/index.html}.}
Condition (iii) is represented by the Hiroshima AI Process launched under
Japan's G7 presidency in 2023.\footnote{See OECD, ``G7 Hiroshima Process on
Generative Artificial Intelligence (AI)'',
\url{https://www.oecd.org/en/publications/g7-hiroshima-process-on-generative-artificial-intelligence-ai_bf3c0c60-en.html};
JapanGov, ``The Hiroshima AI Process'',
\url{https://www.japan.go.jp/kizuna/2024/02/hiroshima_ai_process.html}.}

This is a conditioned structural uniqueness claim about the conjunction of
these three conditions under the external premises admitted in the Lean file.
It is not a moral exclusion of other cities with histories of war or
technological harm.

\section*{Civilizational Layer and Conditioned Uniqueness}

The Lean file introduces an abstract typeclass
\texttt{CivilizationalResponsibilityTech}.  A technology belongs to this class
when it has real witnesses for two constraints: an inadmissible path in the
responsibility category $R$, and civilizational-scale consequences through such
a path.  Nuclear weapons satisfy these constraints via
\texttt{hiroshima\_historical\_witness}.  AI systems satisfy the inadmissible
path constraint via \texttt{bypassBeaconPath} and the civilizational-scale
constraint via the explicit axiom
\texttt{ai\_has\_civilizational\_potential}.  The Lean theorem
\[
\texttt{nuclear\_ai\_same\_civilizational\_layer}
\]
therefore formalizes "same layer" as shared membership in an abstract class
with non-trivial witnesses, not as equality of identical Boolean records.

The Lean file also proves \texttt{two\_crts\_share\_R}: any two instances of
\texttt{CivilizationalResponsibilityTech} provide inadmissible-path witnesses
in the same responsibility category $R$.  Thus the relevant universality claim
is not that two separate categories \texttt{R\_nuclear} and \texttt{R\_ai} are
isomorphic.  The stronger and simpler claim is that $R$ is the fixed
responsibility target shared by all CRT instances.  This is not a claim of
moral equivalence and does not claim that AI harm is inevitable.

The Lean file also proves a conditioned uniqueness theorem:
\[
\texttt{hiroshima\_conditioned\_unique}.
\]
The theorem states that any city satisfying both conditions
\[
\text{IntentionalNuclearAttack}(x)
\quad\text{and}\quad
\text{AIGovernanceAnchor}(x)
\]
is equal to Hiroshima.  This is proved from explicit historical axioms:
Hiroshima and Nagasaki received intentional nuclear attacks, only those cities
satisfy that condition, Hiroshima has an AI governance anchor, and Nagasaki
does not have the corresponding anchor in this model.

The theorem does not prove that AI governance requires Hiroshima, nor that
Hiroshima is necessary for global adoption.  It proves only the conditioned
uniqueness claim stated above.

\section*{Four-Stage AI Logistics Chain}

The Lean file proves a four-stage chain for next-generation AI logistics.

First, \texttt{admissible\_autonomous\_logistics\_needs\_assurance} proves that
any admissible next-generation AI logistics path must pass through
Verification $V$.  This is the formal role of AI assurance in the model and
uses no new axiom beyond the already proved
\texttt{not\_passV\_not\_admissible}.

Second, \texttt{hiroshima\_conditioned\_unique} identifies Hiroshima as the
city satisfying the model's conditioned uniqueness premise:
\texttt{IntentionalNuclearAttack} and \texttt{AIGovernanceAnchor}.

Third, \texttt{hiroshima\_unique\_for\_ai\_logistics} extends this uniqueness
statement with \texttt{WorldLogisticsInfrastructure}.  In the Lean file,
\texttt{WorldLogisticsInfrastructure} is not a single axiom.  It is a defined
conjunction of five components: manufacturing hub, port and sea gateway,
maritime heritage, tourism and human-flow infrastructure, and right-sized
operational scale.  The theorem \texttt{hiroshima\_world\_logistics} proves the
conjunction for Hiroshima from the five component premises, including
\texttt{hiroshima\_right\_sized\_scale}.  The infrastructure condition
strengthens the premise; uniqueness still follows from the first two
conditions.

Fourth, \texttt{hiroshima\_poc\_environment} records the external premise that
Hiroshima has the proof-of-concept environment for implementation.

The combined theorem
\[
\texttt{hiroshima\_four\_stage\_ai\_logistics\_chain}
\]
states all four components together, including both
\texttt{WorldLogisticsInfrastructure City.Hiroshima} and
\texttt{PoCEnvironment City.Hiroshima}.  The theorem does not assert an
exclusive physical build location for logistics systems.  The conclusion is
that the
structural basis, conditional uniqueness, logistics infrastructure, and PoC
environment converge in Hiroshima under the explicit historical,
institutional, and logistical premises stated as axioms in the Lean file.  The
strength of this convergence is conditional on those premises; it is not an
unconditional assertion.

\section*{Definitions}

\begin{definition}[Ordinary corporate decision category]
Let $\mathcal{C}$ be the free category generated by
\[
  D \xrightarrow{p} P \xrightarrow{q} O.
\]
Here $D$ is a corporate decision, $P$ is profit/efficiency evaluation, and $O$
is ordinary operational outcome.  The ordinary decision composite is
\[
  u:=q\circ p:D\to O.
\]
\end{definition}

\begin{definition}[Hiroshima responsibility category]
Let $\mathcal{R}$ be the free category generated by
\[
  D_H \xrightarrow{m} P_H \xrightarrow{b} B
  \xrightarrow{e} E \xrightarrow{v} V \xrightarrow{a} A
  \xrightarrow{o} O'.
\]
Here $D_H$ is a Hiroshima-contextualized corporate decision, $P_H$ is a
Hiroshima-contextualized profit/efficiency evaluation, $B$ is the Beacon
admissibility condition, $E$ is evidence, $V$ is verifiable evidence or
third-party verification, $A$ is accountable decision, and $O'$ is accountable
operational outcome.
\end{definition}

\section*{What $D_H$ Does Not Formally Define}

The notation $D_H$ denotes a corporate decision under Hiroshima place-context.
This formalization does not define ``Hiroshima place-context'' internally as a
mathematical object.  Instead, it records the following external premise:

\begin{itemize}
  \item A historical instance structurally interpreted as
  $\neg\mathrm{Adm}_{B,V}$ at civilizational scale exists and is located in
  Hiroshima.
  \item This instance is preserved as institutional world memory.
  \item Hiroshima has already been used as an international AI governance
  reference point through the Hiroshima AI Process.
\end{itemize}

Thus $D_H$ is not arbitrary contextualization.  It is the starting object for
decisions interpreted under a historically documented responsibility context.

\begin{definition}[Hiroshima responsibility functor]
Define the Hiroshima Responsibility Functor
\[
  F_H:\mathcal{C}\to\mathcal{R}
\]
on objects by
\[
  F_H(D)=D_H,\qquad F_H(P)=P_H,\qquad F_H(O)=O'.
\]
Define it on generating morphisms by
\[
  F_H(p)=m:D_H\to P_H,
\]
and
\[
  F_H(q)=o\circ a\circ v\circ e\circ b:P_H\to O'.
\]
Since $\mathcal{C}$ is free on the generators $p,q$, this uniquely extends to
a functor on all morphisms of $\mathcal{C}$.
\end{definition}

\begin{definition}[Admissibility through Beacon and Verification]
For a morphism $f:X\to Y$ in $\mathcal{R}$, define
\[
  f\in \mathrm{Adm}_{B,V}(X,Y)
\]
if and only if there exist morphisms
\[
  x:X\to B,\qquad y:B\to V,\qquad z:V\to Y
\]
such that
\[
  f=z\circ y\circ x.
\]
Equivalently, $f$ factors through both Beacon $B$ and Verification $V$.
\end{definition}

\begin{remark}
In the Lean formalization, $\mathrm{Adm}_{B,V}$ is implemented via
\texttt{pathObjects}, which records the sequence of objects visited by a path.
The equivalence between this implementation and the factorization definition
above is proved as \texttt{AdmBV\_iff\_factorizes} in the Lean file.  The key
auxiliary is a path-splitting lemma: if an object appears in
\texttt{pathObjects}$\,p$, then $p$ splits at that object.
\end{remark}

\section*{Propositions}

\begin{proposition}[Functorial image of ordinary decision]
The Hiroshima responsibility functor sends the ordinary decision composite
$u=q\circ p$ to the responsibility composite
\[
  F_H(u)
  =
  o\circ a\circ v\circ e\circ b\circ m
  :
  D_H\to O'.
\]
\end{proposition}

\begin{proof}
By functoriality,
\[
  F_H(q\circ p)=F_H(q)\circ F_H(p).
\]
Substituting the definitions gives
\[
  F_H(q)\circ F_H(p)
  =
  (o\circ a\circ v\circ e\circ b)\circ m.
\]
By associativity of composition, this is
\[
  o\circ a\circ v\circ e\circ b\circ m:D_H\to O'.
\]
\end{proof}

\begin{proposition}[Responsibility composite is admissible]
The morphism
\[
  o\circ a\circ v\circ e\circ b\circ m:D_H\to O'
\]
belongs to $\mathrm{Adm}_{B,V}(D_H,O')$.
\end{proposition}

\begin{proof}
Let
\[
  x=b\circ m:D_H\to B,\qquad
  y=v\circ e:B\to V,\qquad
  z=o\circ a:V\to O'.
\]
Then
\[
  z\circ y\circ x
  =
  (o\circ a)\circ(v\circ e)\circ(b\circ m)
  =
  o\circ a\circ v\circ e\circ b\circ m.
\]
Hence the responsibility composite factors through both $B$ and $V$.
\end{proof}

\begin{proposition}[Beacon bypass is inadmissible]
For any path $f:X\to Y$ in $\mathcal{R}$, if $f$ does not factor through $B$,
then $f\notin \mathrm{Adm}_{B,V}(X,Y)$.
\end{proposition}

\begin{proof}
By definition, $f\in\mathrm{Adm}_{B,V}(X,Y)$ only if there exist
$x:X\to B$, $y:B\to V$, and $z:V\to Y$ such that
$f=z\circ y\circ x$.  This is a factorization through $B$.  Therefore, if no
such factorization through $B$ exists, $f$ is not admissible.
\end{proof}

\begin{proposition}[Verification bypass is inadmissible]
For any path $f:X\to Y$ in $\mathcal{R}$, if $f$ does not factor through $V$,
then $f\notin \mathrm{Adm}_{B,V}(X,Y)$.
\end{proposition}

\begin{proof}
By definition, $f\in\mathrm{Adm}_{B,V}(X,Y)$ only if there exist
$x:X\to B$, $y:B\to V$, and $z:V\to Y$ such that
$f=z\circ y\circ x$.  This is a factorization through $V$.  Therefore, if no
such factorization through $V$ exists, $f$ is not admissible.
\end{proof}

\section*{Diagram}

\[
\begin{tikzcd}[column sep=large,row sep=large]
\mathcal{C}:
  D \arrow[r,"p"] & P \arrow[r,"q"] & O
\\[1.5em]
\mathcal{R}:
  D_H \arrow[r,"m"] &
  P_H \arrow[r,"b"] &
  B \arrow[r,"e"] &
  E \arrow[r,"v\;(\mathrm{ADIC})"] &
  V \arrow[r,"a"] &
  A \arrow[r,"o"] &
  O'
\end{tikzcd}
\]

The functor $F_H$ maps $D,P,O$ to $D_H,P_H,O'$ and maps the ordinary evaluation
process to the responsibility process.

\section*{Compact Interpretation}

Hiroshima placement is a functor from the ordinary corporate decision category
$\mathcal{C}$ to the Hiroshima responsibility category $\mathcal{R}$:
\[
  F_H:\mathcal{C}\to\mathcal{R}.
\]
It maps corporate decision $D$, profit/efficiency evaluation $P$, and outcome
$O$ to Hiroshima-contextualized decision $D_H$, Hiroshima-contextualized
profit/efficiency evaluation $P_H$, and accountable outcome $O'$.

It sends the ordinary composite
\[
  D\to P\to O
\]
to the responsibility composite
\[
  D_H\to P_H\to B\to E\to V\to A\to O'.
\]
Therefore, a path that bypasses Beacon $B$ or Verification $V$ is not an
admissible responsible morphism under $\mathrm{Adm}_{B,V}$.

\section*{What Is and Is Not Claimed}

\begin{itemize}
  \item Claimed: $F_H$ is a functor from the ordinary decision category to the
  Hiroshima responsibility category.
  \item Claimed: $F_H(q\circ p)=o\circ a\circ v\circ e\circ b\circ m$.
  \item Claimed: this image factors through both Beacon $B$ and Verification
  $V$.
  \item Claimed: paths bypassing $B$ or $V$ are not in
  $\mathrm{Adm}_{B,V}$.
  \item Claimed as an external historical premise: under the admitted
  historical and institutional premises, Hiroshima is the city satisfying the
  conjunction of (i) nuclear-layer
  $\neg\mathrm{Adm}_{B,V}$ witness, (ii) institutional preservation as world
  memory, and (iii) existing AI governance anchor through the Hiroshima AI
  Process.
  \item Claimed as a Lean theorem: under the predicates
  \texttt{IntentionalNuclearAttack} and \texttt{AIGovernanceAnchor},
  Hiroshima is selected by the conditioned uniqueness result.
  \item Claimed \emph{conditionally}: under the explicit premises stated as
  axioms in the Lean file, the responsibility structure for next-generation AI
  logistics most naturally converges to Hiroshima.  The convergence is as
  strong as the premises admitted.  It is not an unconditional assertion.
  \item Not claimed: Hiroshima uniquely guarantees ethical action.
  \item Not claimed: AI governance requires Hiroshima as an anchor, or that
  Hiroshima is necessary for global adoption.
  \item Not claimed: that Hiroshima is unconditionally or uniquely necessary
  for world AI logistics development.  The uniqueness result
  (\texttt{hiroshima\_conditioned\_unique}) holds under the stated two
  conditions; it does not hold independently of those conditions.
  \item Not claimed: autonomous AI logistics systems cannot be physically built
  or operated outside Hiroshima.  The claim concerns the structural origin of
  the responsibility architecture, assurance anchor, logistics infrastructure,
  and PoC environment.
  \item Not claimed: other cities with war histories are excluded on moral
  grounds.  They are excluded from this specific role only if they lack the
  conjunction required above.
  \item Not claimed: psychological pressure itself is mathematically proved.
  \item Not claimed: category theory alone proves institutional compliance.
\end{itemize}

\end{document}
